\begin{center}
\textit{Tune: Outward Bound}
\end{center}

\settowidth{\versewidth}{To \textit{Suke} \& \textit{Sall} \& \textit{Kitty} too.}
\begin{verse}[\versewidth]
To the \textsc{Liverpool} docks we bid adieu,\\*
To \textit{Suke} \& \textit{Sall} \& \textit{Kitty} too.\\
Our anchor's wieghed and our sails unfurled;\\*
We're off to plough the watery world.\\!

{\itshape 
Hurrah! We're outward bound!\\*
Hurrah! We're outward bound!}\\!

When the wind it blows from the east-nor'-east,\\*
Our ship will sail 10\textsc{kts} at least.\\
The purser will our wants supply,\\*
So while we've rum we'll never say die.\\!

And should we touch at \textsc{Malabar}\\*
Or any other port as far,\\
Our purser he will tip the chink\\*
And just like fishes we will drink.\\!

One day the man on the look-out\\*
Proclaims a sail with a joyful shout:\\
`Can you make her out?' `I think I can.\\*
She's a pilot standing out from the land.'\\!

{\itshape 
Hurrah! We're homeward bound!\\*
Hurrah! We're homeward bound!}\\!

Now when we get to the \textsc{Blackwall} docks,\\*
The pretty young girls come down in flocks;\\
One to the other you'll hear them say,\\*
`O here comes \textit{Jack} with his 10 months' pay.\\!

And when we get to the \textsc{Dog and Bell},\\*
It's there they've got good liquor to sell.\\
In comes old \textit{Grouse} with a smile,\\*
Saying, `Drink, my boys. It's worth your while.'\\!

But when the money's all gone \& spent,\\*
And there's none to be borrowed and none to be lent,\\
In comes old \textit{Grouse} with a frown,\\*
Saying, `Get up, \textit{Jack}. Let \textit{John} sit down.\\!

Then poor old \textit{Jack} must understand\\*
There's ships in docks all wanting hands;\\
So he goes onboard as he did before,\\*
And bids adieu to his native shore.\\!

{\itshape 
Hurrah! We're outward bound!\\*
Hurrah! We're outward bound!}
\end{verse}