\documentclass[MAIN]{subfiles}
\begin{document}

\settowidth{\versewidth}{My heart aches, and a drowsy numbness pains}
\begin{verse}[\versewidth]
\poemlines{5}
My heart aches, and a drowsy numbness pains\footnotetext{$\dagger$`Ode to a Nightingale', John Keats (1795 -- 1821). \cite{treasury}. \P Where the Almanackist gives `foreign', Palgrave gives `alien'; `foreign' is the Almanackist's invention, but `alien' is an intolerable metrical sin.}\\*
\vin My sense, as though of hemlock I had drunk,\\
Or emptied some dull opiate to the drains\\
\vin One minute past, and {\sc Lethe}-wards had sunk:\\
'Tis not through envy of thy happy lot,\\
\vin But being too happy in thine happiness --\\
\vin \vin That thou, light-wing\`ed dryad of the trees\\
\vin \vin \vin \vin In some melodious plot\\
\vin Of beechen green, \& shadows numberless,\\*
\vin \vin Singest of summer in full-throated ease.\\!

O for a draught of vintage! that hath been\\*
\vin Cooled a long age in the deep-delv\`ed earth,\\
Tasting of \emph{Flora} \& the country green,\\
\vin Dance, \& provencal song, \& sunburnt mirth!\\
O for a beaker full of the warm south,\\
\vin Full of the true, the blushful {\sc Hippocrene},\\
\vin \vin With beaded bubbles winking at the brim,\\
\vin \vin \vin \vin And purple-stain\`ed mouth;\\
\vin That I might drink, and leave the world unseen,\\*
\vin \vin And with thee fade away into the forest dim:\\!

Fade far away, dissolve, and quite forget\\*
\vin What thou among the leaves hast never known,\\
The weariness, the fever, \& the fret\\
\vin Here, where men sit and hear each other groan;\\
Where palsy shakes a few, sad, last gray hairs,\\
\vin Where youth grows pale, \& spectre-thin, and dies;\\
\vin \vin Where but to think is to be full of sorrow\\
\vin \vin \vin \vin And leaden-eyed despairs,\\
\vin Where beauty cannot keep her lustrous eyes,\\*
\vin \vin Or new Love pine at them beyond tomorrow.\\!

Away! away! for I will fly to thee,\\*
\vin Not charioted by \emph{Bacchus} \& his pards,\\
But on the viewless wings of poesy,\\
\vin Though the dull brain perplexes \& retards:\\
Already with thee! tender is the night,\\
\vin And haply the queen-moon is on her throne,\\
\vin \vin Clustered around by all her starry fays;\\
\vin \vin \vin \vin But here there is no light,\\
\vin Save what from heaven is with the breezes blown\\*
\vin \vin Through verdurous glooms \& winding mossy ways.\\!

I cannot see what flowers are at my feet,\\*
\vin Nor what soft incense hangs upon the boughs,\\
But, in embalm\`ed darkness, guess each sweet\\
\vin Wherewith the seasonable month endows\\
The grass, the thicket, \& the fruit-tree wild;\\
\vin White hawthorn, \& the pastoral eglantine;\\
\vin \vin Fast fading violets covered up in leaves;\\
\vin \vin \vin \vin And mid-may's eldest child,\\
\vin The coming musk-rose, full of dewy wine,\\*
\vin \vin The murmurous haunt of flies on summer eves.\\!

Darkling I listen; and, for many a time\\*
\vin I have been \sfrac{$1$}{$2$} in love with easeful death,\\
Called him soft names in many a mus\`ed rhyme,\\
\vin To take into the air my quiet breath;\\
Now more than ever seems it rich to die,\\
\vin To cease upon the midnight with no pain,\\
\vin \vin While thou art pouring forth thy soul abroad\\
\vin \vin \vin \vin In such an ecstasy!\\
\vin Still wouldst thou sing, and I have ears in vain --\\*
\vin \vin To thy high requiem become a sod.\\!

Thou wast not born for death, immortal bird!\\
\vin No hungry generations tread thee down;\\
The voice I hear this passing night was heard\\
\vin In ancient days by emperor \& clown:\\
Perhaps the selfsame song that found a path\\
\vin Through the sad heart of \emph{Ruth}, when, sick for home,\\
\vin \vin She stood in tears amid the foreign corn;\\
\vin \vin \vin \vin The same that oft-times hath\\
\vin Charmed magic casements, opening on the foam\\*
\vin \vin Of perilous seas, in fairy lands forlorn.\\!

Forlorn! The very word is like a bell\\*
\vin To toll me back from thee to my sole self!\\
Adieu! The fancy cannot cheat so well\\
\vin As she is famed to do, deceiving elf.\\
Adieu! Adieu! Thy plaintive anthem fades\\
\vin Past the near meadows, over the still stream,\\
\vin \vin Up the hillside; and now 'tis buried deep\\
\vin \vin \vin \vin In the next valley-glades:\\
\vin Was it a vision, or a waking dream?\\*
\vin \vin Fled is that music. Do I wake or sleep?
\end{verse}

\end{document}