\documentclass[MAIN]{subfiles}
\begin{document}

\settowidth{\versewidth}{So over seven-acre field \& up-along across the down}
\begin{verse}[\versewidth]
\poemlines{5}
\vin Three summers since I chose a maid,\footnotetext{$\xi$ `The Farmer's Bride', Miss Charlotte Mew (1869 -- 1928). \cite{norton}. Miss Mew never married, and was fond of wearing men's clothes.}\\*
\vin Too young maybe -- but more's to do\\
\vin At harvest-time than bide \& woo.\\
\vin \vin When us was wed she turned afraid\\
\vin Of love \& me \& all things human;\\
\vin Like the shut of a winter's day\\
\vin Her smile went out, and 'twasn't a woman --\\
\vin More like a little frightened fay.\\*
\vin \vin One night, in the fall, she runned away.\\!

\vin `Out 'mong the sheep, her be,' they said;\\*
\vin 'Should properly have been abed;\\
\vin But sure enough she wasn't there\\
\vin Lying awake with her wide brown stare.\\
So over seven-acre field \& up-along across the down\\
\vin We chased her, flying like a hare\\
\vin Before out lanterns. To {\sc Church-Town}\\
\vin \vin All in a shiver \& a scare\\
\vin We caught her, fetched her home at last\\*
\vin \vin And turned the key upon her, fast.\\!

\vin She does the work about the house\\*
\vin As well as most, but like a mouse:\\
\vin Happy enough to chat \& play\\
\vin \vin With birds \& rabbits \& such as they,\\
\vin \vin So long as men-folk keep away.\\
\vin `Not near, not near!' her eyes beseech\\
\vin When one of us comes within reach.\\
\vin \vin The women say that beasts in stall\\
\vin \vin Look round like children at her call.\\*
\vin \vin I've hardly heard her speak at all.\\!

\vin Shy as a leveret, swift as he,\\*
\vin Straight \& slight as a young larch tree,\\
\vin Sweet as the first wild violets, she,\\*
\vin To her wild self. But what to me?\\!

\vin The short days shorten \& the oaks are brown;\\*
\vin \vin The blue smoke rises to the low grey sky;\\
\vin One leaf in the still air falls slowly down;\\
\vin \vin A magpie's spotted feathers lie\\
\vin On the black earth spread white with rime;\\
\vin The berries redden up to Christmas-time.\\
\vin \vin What's Christmas-time without there be\\*
\vin \vin Some other in the house than we!\\!

\vin \vin She sleeps up in the attic there\\*
\vin \vin Alone, poor maid. 'Tis but a stair\\
\vin Betwixt us. Oh! my God! the down,\\
\vin The soft young down of her, the brown,\\*
The brown of her -- her eyes, her hair, her hair!
\end{verse}

\end{document}