\documentclass[MAIN]{subfiles}
\begin{document}

\setarab
\vocalize
\arabtrue

\settowidth{\versewidth}{\vin From the topaz mountain to the desert of doubt,}
\begin{verse}[\versewidth]
\poemlines{5}
Who are these from the strange ineffable places,\footnotetext{$\zeta$ `Arabia', John Falkner (1858 -- 1932). \cite{oxfordlarkin}. The poem has the subtitle: `[David George] Hogarth's \emph{Penetration of Arabia}'. \P This poem contains a number of intriguingly obscure references, beginning with a roll-call of significant -- though now largely forgotten -- European explorers in the third verse. Jean Louis Burckhardt (1784 -- 1817) was a Swiss explorer and the first European to set eyes on the city of Petra in over a thousand years. Joseph Hal\'evy (1827 -- 1917) was an Ottoman-French-Jewish orientalist who was most notable for his exploration of the Yemen. Karsten Niebuhr (1733 -- 1815) was the cartographer of the Royal Danish Arabian Expedition, and the only member of that group to return to Europe alive. Ulrich Jasper Seetzen (1767 -- 1811) was murdered as an infidel by his fellow Muslims -- he had converted two years before -- while in search of the lost city that Burckhardt would finally rediscover. George Sadleir (1789 -- 1859) was a captain in the British Army who, in endeavouring (successfully) to deliver a ceremonial sword to an Egyptian commander on behalf of Queen Victoria, inadvertently became the first European to cross the Arabia Peninsula. Jan Jansz Struys (1630 -- 1694) was a Dutch sailor more famous for exploring Russia, but who, as prisoner of war in the Ottoman Empire, must have seen more of the Middle East than most Europeans of his day. The exact Slater being referred to, however, remains unclear. \P The location of Samna is likewise unclear. Is this perhaps an archaic name for -- or a garbled version of -- Sana'a? \P Zob\"eide is an archaic romanisation of \RL{zbydT}, now more commonly transliterated as Zubaidah, the granddaughter, niece and wife of three distinct Abbasid caliphs, famous for constructing a series of aqueducts for Mecca and Medina.}\\*
\vin From the topaz mountain to the desert of doubt,\\
With the glow of the Yemen full on their faces,\\*
\vin And a breath from the spices of Hadramaut?\\!

Travel-apprentices, travel-indenturers,\\*
\vin Young men, old men, black hair, white,\\
Names to conjure with, wild adventurers,\\*
\vin From the noonday furnace to the purple night.\\!

\emph{Burckhardt}, \emph{Hal\'evy}, \emph{Niebuhr}, \emph{Slater},\\*
\vin Seventeenth, 18th century beys,\\
\emph{Seetzen}, \emph{Sadleir}, \emph{Struys} and later\\*
\vin Down to the long victorian days.\\!

A 1000 miles at the back of {\sc Aden},\\*
\vin There they had time to think of things;\\
In the outer silence and burnt air laden\\*
\vin With the shadow of death \& a vulture's wings.\\!

There they remembered the last house in {\sc Samna},\\*
\vin Last of the plane-trees, last shepherd \& flock,\\
Prayed for the heavens to rain down manna,\\*
\vin Prayed for a \emph{Moses} to strike down the rock.\\!

Famine \& fever flagged their forces\\*
\vin Till they died in a dream of ice \& fruit\\
In the long-forgotten watercourses\\*
\vin By the edge of Queen \emph{Zob\"eide}'s route.\\!

They have left the hope of the green oases,\\*
\vin The fear of the bleaching bones \& the pest,\\
They have found the more ineffable places --\\*
\vin {\hge Allah} has given them rest.
\end{verse}

\end{document}
