\documentclass[MAIN]{subfiles}
\begin{document}

\settowidth{\versewidth}{`I thought it was love you meant.' `Well, it's done.' `Yes, it's done.}
\begin{verse}[\versewidth]
\poemlines{5}
`I thought you loved me.' `No, it was only fun.'\footnotetext{`In the Orchard', Mrs Muriel Board (1885 -- 1967). \cite{oxfordlarkin}. Mrs Board tended to publish her work under her maiden name Stuart.}\\*
`When we stood there, closer than all?' `Well, the harvest moon\\
Was shining \& queer in your hair, and it turned my head.'\\
`That made you?' `Yes.' `Just the moon \& the light it made\\
Under the tree?' `Well, your mouth, too.' `Yes, my mouth?'\\
`And the quiet there that sang like the drum in the booth.\\
You shouldn't have danced like that.' `Like what?' `So close,\\
With your head turned up, and the flower in your hair, a rose\\
That smelt all warm.' `I loved you. I thought you knew\\
I wouldn't have danced like that with any but you.'\\
`I didn't know, I thought you knew it was fun.'\\
`I thought it was love you meant.' `Well, it's done.' `Yes, it's done.\\
I've seen boys stone a blackbird, and watched them drown\\
A kitten... It clawed at the reeds, and they pushed it down\\
Into the pool while it screamed. Is that fun, too?'\\
`Well, boys are like that... Your brothers...' `Yes, I know.\\
But you, so lovely \& strong! Not you! Not you!'\\
`They don't understand it's cruel. It's only a game.'\\
`And are girls fun, too?' `No, still in a way it's the same.\\
It's queer \& lovely to have a girl...' `Go on.'\\
`It makes you mad for a bit to feel she's your own,\\
And you laugh and kiss her, and maybe you give her a ring,\\
But it's only in fun.' `But I gave you everything.'\\
`Well, you shouldn't have done it. You know what a fellow thinks\\
When a girl does that.' `Yes, he talks of her over his drinks\\
And calles her a --' `Stop that now, I thought you knew.'\\
`But it wasn't with anyone else. It was only you.'\\
`How did I know? I thought you wanted it too.\\
I thought you were like the rest. Well, what's to be done?'\\
`To be done' `Is it all right?' `Yes.' `Sure?' `Yes, but why?'\\
`I don't know; I thought you were going to cry.\\
You said you had something to tell me.' `Yes, I know.\\
It wasn't anything really... I think I'll go.'\\
`Yes, it's late. There's thunder about, a drop of rain\\
Fell on my hand in the dark. I'll see you again\\
At the dance next week. You're sure that everything's right?'\\*
`Yes,' `Well, I'll be going.' `Kiss me...' `Good night...' `Good night.'
\end{verse}

\end{document}