\documentclass[MAIN]{subfiles}
\begin{document}

\settowidth{\versewidth}{My pensive \emph{Sara}, thy soft cheek reclined}
\begin{verse}[\versewidth]
\poemlines{5}
My pensive \emph{Sara}, thy soft cheek reclined\footnotetext{`The Eolian Harp', Samuel Coleridge (1772 -- 1834). \cite{norton}. \emph{Eolian} is the spelling of the \emph{OED}'s \emph{Aeolian} preferred by Dr Johnson, whose \emph{Dictionary} Coleridge himself appears to have used; an Aeolian harp is a kind of stringed instrument, designed so that it can be played entirely by the wind. Coleridge's own note indicates that this poem was written at Clevedon in Somerset.}\\*
Thus on mine arm, most soothing sweet it is\\*
To sit beside our cot, our cot o'ergrown\\
With white-flowered jasmine, and the broad-leaved myrtle,\\ 
(Meet emblems they of innocence and love!)\\
And watch the clouds, that late were rich with light,\\
Slow saddening round, and mark the star of eve\\
Serenely brilliant (such would wisdom be)\\
Shine opposite! How exquisite the scents\\
Snatched from yon bean-field! And the world s\'o hushed!\\
The stilly murmur of the distant sea\\
Tells us of silence. And that simplest lute,\\
Placed length-ways in the clasping casement, hark!\\
How by the desultory breeze caressed,\\
Like some coy maid \sfrac{$1$}{$2$} yielding to her lover,\\
It pours such sweet upbraiding, as must needs\\
Tempt to repeat the wrong! And now, its strings\\
Boldlier swept, the long sequacious notes\\
Over delicious surges sink \& rise,\\
Such a soft floating witchery of sound\\
As twilight elfins make, when they at eve\\
Voyage on gentle gales from fairy land,\\
Where melodies round honey-dropping flowers,\\
Footless and wild, like birds of paradise,\\
Nor pause, nor perch, hovering on untamed wing!\\
O the one life within us and abroad,\\
Which meets all motion and becomes its soul,\\
A light in sound, a sound-like power in light,\\
Rhythm in all thought, and joyance everywhere --\\
Methinks, it should have been impossible\\
Not to love all things in a world so filled;\\
Where the breeze warbles, and the mute still air\\*
Is music slumbering on her instrument.\\!

And thus, my love! as on the midway slope\\* 
Of yonder hill I stretch my limbs at noon,\\
Whilst through my half-closed eyelids I behold\\ 
The sunbeams dance, like diamonds, on the main,\\
And tranquil muse upon tranquility:\\
Full many a thought uncalled and undetained,\\
And many idle flitting phantasies,\\
Traverse my indolent \& passive brain,\\
As wild and various as the random gales\\*
That swell \& flutter on this subject lute!\\!

And what if all of animated nature\\*
Be but organic harps diversely framed,\\
That tremble into thought, as o'er them sweeps\\ 
Plastic and vast, one intellectual breeze,\\*
At once the soul of each, and God of all?\\!

But thy more serious eye a mild reproof\\*
Darts, O belov\`ed woman! nor such thoughts\\
Dim \& unhallowed dost thou not reject,\\
And biddest me walk humbly with my God.\\
Meek daughter in the family of \emph{Christ}!\\
Well hast thou said and holily dispraised\\
These shapings of the unregenerate mind;\\
Bubbles that glitter as they rise and break\\
On vain philosophy's aye-babbling spring.\\ 
For never guiltless may I speak of him,\\
The incomprehensible! save when with awe\\
I praise him, and with faith that inly f\'eels;\\
Who with his saving mercies heal\`ed me,\\
A sinful \& most miserable man,\\ 
Wildered \& dark, and gave me to possess\\*
Peace, and this cot, and thee, heart-honored maid! 
\end{verse}

\end{document}