\documentclass[MAIN]{subfiles}
\begin{document}

\settowidth{\versewidth}{When they first reached the dock they scarce could stand,}
\begin{verse}[\versewidth]
\poemlines{5}
'Twas going to snow. 'Twas snowing. Curse his luck.\footnotetext{$\ddagger$ `The Drove-Road', Wilfrid Gibson (1878 -- 1962). \cite{oxfordlarkin}. \P The Irish Channel is an archaic name for the North Channel which connects the Irish Sea with the Atlantic.}\\*
And 15 mile to travel. Here was he\\
With nothing but an empty pipe to suck,\\
And \sfrac{$1$}{$2$} a flask of rum -- but that would be\\
More welcome later on. He'd had a drink\\
Before he left, and that would keep him warm\\
A tidy while; and 'twould be good to think\\
He'd something to fall back on if the storm\\
Should come to much. You never knew with snow.\\
A sup of rain he didn't mind at all,\\
But snow was different with so far to go --\\
Full 15 mile, and not a house of call.\\
Ay, snow was quite another story, quite --\\
Snow on the fell-tops with a north-east wind\\
Behind it, blowing steadily with a bite\\
That made you feel that you were stark \& skinned.\\
And those poor beasts -- and they just off the boat\\
A day or so, and hardly used to land --\\
Still dizzy with the sea, their wits afloat.\\
When they first reached the dock they scarce could stand,\\
They'd been so joggled. It's gey bad to cross,\\
After a long day's jolting in the train,\\
Thon Irish Channel, always pitch \& toss --\\
And, heads or tails, not much for them to gain.\\
And then the market, and the throng \& noise\\
Of yapping dogs; and they stung mad with fear,\\
Welted with switches by those senseless boys --\\
He'd like to dust their jackets. But 'twas queer,\\
A beast's life, when you came to think of it,\\
From start to finish -- queerer, ay, a lot\\
Than any man's, and chancier a good bit.\\
With his ash-sapling at their heels they'd got\\
To travel before night those 15 miles\\
Of hard fell road against the driving snow,\\
Half-blinded, on \& on. He thought at whiles\\
'Twas just as well for them they couldn't know...\\
Though, as for that, 'twas little that he knew\\
Himself what was in store for him. He took\\
Things as they came: 'twas all a man could do;\\
And he'd kept going somehow by hook or crook.\\
And here he was, with 15 mile of fell,\\
And snow and... God, but it was blowing stiff.\\
And no tobacco. Blessed if he could tell\\
Where he had lost it -- but for \sfrac{$1$}{$2$} a whiff\\
He'd swop the very jacket off his back --\\
Not that he'd miss the cobweb of old shreds\\
That held the holes together. Thon cheap-jack\\
Who'd sold it him had said it was Lord \emph{Ted}'s,\\
And london-cut: but \emph{Teddy} had grown fat\\
Since he'd been made an alderman... His bid?\\
And did the gentleman not want a hat\\
To go with it, a topper? If he did,\\
Here was the very... Hell, but it was cold,\\
And driving dark it was -- nigh dark as night.\\
He'd almost think he must be getting old\\
To feel the wind so. And long out of sight\\
The beasts had trotted. Well, what odds. The way\\
Ran straight for 10 miles on, and they'd go straight:\\
They'd never heed a by-road. Many a day\\
He'd had to trudge on, trusting them to fate,\\
And always found them safe. They scamper fast,\\
But in the end a man could walk them down:\\
They're showy trotters, but they cannot last:\\
He'd race the fastest beast for half-a-crown\\
On a day's journey. Beasts were never made\\
For steady travelling -- drive them 20 mile\\
And they were done; while he was not afraid\\
To travel twice that distance with a smile.\\
But not a day like this. He'd never felt\\
A wind with such an edge. 'Twas like the blade\\
Of the rasper in the pocket of his belt\\
He kept for easy shaving. In his trade\\
You'd oft to make your toilet under a dyke --\\
And he was always one for a clean chin,\\
And carried soap. He'd never felt the like --\\
That wind, it cut clean through you to the skin.\\
He might be mother-naked, walking bare,\\
For all the use his clothes were, with the snow\\
Half blinding him and clagging to his hair\\
And trickling down his spine. He'd like to know\\
What was the sense of pegging steadily,\\
Chilled to the marrow, after a daft herd\\
Of draggled beasts he couldn't even see.\\
But that was him all over -- just a word,\\
A nod, a wink, the price of \sfrac{$1$}{$2$} \& \sfrac{$1$}{$2$},\\
And he'd be setting out for God-knows-where\\
With no more notion than a yearling calf\\
Where he would find himself when he got there.\\
And he'd been travelling hard on 60 year\\
The same old road, the same old giddy gait;\\
And he'd be walking, for a pint of beer,\\
Into his coffin one day, soon or late --\\
But not with such a tempest in his teeth,\\
Half-blinded and \sfrac{$1$}{$2$} dothered, that he hoped.\\
He'd met a sight of weather on the heath,\\
But this beat all. 'Twas worse than when he'd groped\\
His way that evening down the {\sc Mallerstang} --\\
Thon was a blizzard, thon, and he was done\\
And almost dropping, when he came a bang\\
Against a house -- slap-bang, and like to stun.\\
Though that just saved his senses: and right there\\
He saw a lighted window he'd not seen,\\
Although he'd nearly staggered through its glare\\
Into a goodwife's kitchen, where she'd been\\
Baking hot griddle-cakes upon the peat...\\
And he could taste them now, and feel the glow\\
Of steady, aching, tingly, drowsy heat\\
As he sat there and let the caking snow\\
Melt off his boots, staining the sanded floor.\\
And that brown jug she took down from the shelf --\\
And every time he'd finished fetching more\\
And piping, `Now reach up and help yourself!'\\
She was a wonder, thon, the gay old wife --\\
But no such luck this journey. Things like that\\
Could hardly happen every day of life,\\
Or no one would be dying but the fat\\
And oily undertakers, starved to death\\
For want of custom... Hell, but he would soon\\
Be giving them a job... It caught your breath,\\
That throttling wind. And it was not yet noon.\\
And he'd be travelling through it until dark.\\
Dark! 'Twas already dark, and might be night\\
For all that he could see. And not a spark\\
Of comfort for him. Just to strike a light\\
And press the kindling shag down in the bowl,\\
Keeping the flame well shielded with his hand,\\
And puff \& puff! He'd give his very soul\\
For \sfrac{$1$}{$2$} a pipe. He couldn't understand\\
How he had come to lose it. He'd the rum --\\
Ay, that was safe enough, but it would keep\\
A while: you never knew what chance might come\\
In such a storm... If he could only sleep...\\
If he could only sleep... That rustling sound\\
Of drifting snow, it made him sleepy-like --\\
Drowsy \& dizzy, dithering round \& round...\\
If he could only curl up under a dyke\\
And sleep \& sleep. ... It dazzled him, that white,\\
Drifting \& drifting round \& round \& round...\\
Just \sfrac{$1$}{$2$} a moment's snooze... He'd be all right.\\
It made his head quite dizzy, that dry sound\\
Of rustling snow: it made his head go round,\\
That rustling in his ears... and drifting, drifting...\\
If only he could sleep... He would sleep sound...\\
God, he was nearly gone... The storm was lifting;\\
And he'd run into something soft \& warm --\\
Slap into his own beasts, and never knew.\\
Huddled they were, bamboozled by the storm --\\
And little wonder either when it blew\\
A blasted blizzard. Still, they'd got to go:\\
They couldn't stand there snoozing until night.\\
But they were sniffing something in the snow:\\
'Twas that had stopped them, something big \& white --\\
A bundle -- nay, a woman... And she slept --\\
But it was death to sleep. He'd nearly dropped\\
Asleep himself. 'Twas well that he had kept\\
That rum, and lucky that the beasts had stopped.\\
Ay, it was well that he had kept the rum:\\
He liked his drink, but he had never cared\\
For soaking by himself \& sitting mum:\\*
Even the best rum tasted better shared.
\end{verse}

\end{document}