\documentclass[MAIN]{subfiles}
\begin{document}

\bigskip

\begin{center}
{\it Tune: Cadgwith Anthem}
\end{center}

\bigskip

\settowidth{\versewidth}{We come from yonder mountains. Our pistols are loaded,}
\begin{verse}[\versewidth]
\poemlines{5}
We come from yonder mountains. Our pistols are loaded,\footnotetext{Anon. \cite{herringbone}. George Goodenough wrote in \emph{The Handy Man Afloat and Ashore} (1901): `The following song always struck me as having one of the finest airs ever sung on a foc's'le. To hear the chorus pealing forth from some hundred or more throats was a thing to be remembered. The only pity is that the words are not more sensible. Such as they are they were very difficult to obtain. A bluejacket once wrote down all he could remember of them for me, but the copy got mixed up with other papers and I thought I had lost it. No one else could I find that could repair the supposed loss. Inquiries at second-hand music shops in London were fruitless. Many men could tell me that they knew the song but could not give me the words. Quite recently I came across my copy and here is the song.'}\\*
For to rob and to plunder it is our intent.\\!

{\it As we roam through the valleys\\*
Where the lilies and the roses,\\
And the beautiful cashmere lies drooping its head:\\
Then away, then away, then away, away,\\
To the caves in yonder mountains,\\*
To the robbers' retreat!}\\!

Hark, hark! In the distance there's footsteps approaching.\\*
Stand, stand and deliver, shall be our watchword.\\!

Your gold and your jewels -- your life if resisted!\\*
We shall laugh at your agony, and scorn at your threats.\\!

Come fill up your glasses and let be a-drinking,\\*
For the moonbeams are shining all over our heads.
\end{verse}

\end{document}
