\documentclass[MAIN]{subfiles}

\begin{document}

\begin{center}
\emph{Tune: The Blaydon Races}
\end{center}

\bigskip

\settowidth{\versewidth}{\it I went to {\sc Blaydon} races. 'Twas on the ninth o' june, qqqq}
\begin{verse}[\versewidth]
I went to {\sc Blaydon} races. 'Twas on the ninth o' june,\footnotetext{`The Blaydon Races', George Ridley (1835 -- 1864). \cite{spinners}. The words here are from The Spinners' version; they seem to have modifed Ridley's lyrics to improve the meter.}\\*
Eighteen hundred \& 62 on a summer's afternoon.\\
I took the bus from {\sc Balmbras}, an' she was heavy laden.\\*
Away we went along {\sc Collingwood Street} that's on the road to {\sc Blaydon}.\\!

{\it O lads! You should've seen us gannin'!\\*
Passin' the folks along the road, just as they were stan'in':\\
Al' the lads \& lasses there, an' al' wi' smilin' faces,\\*
Gannin' along the {\sc Scotswood Road} to see the {\sc Blaydon} races.\\!}

We flew past Armstrong's factory an' up to the {\sc Robin Adair}.\\*
Gannin' down the railway bridge, the bus wheel flew off there.\\
The lasses lost their crinolines \& the veils that hide their faces.\\*
I got two black eyes \& a broken nose a-gannin to {\sc Blaydon} races.\\!

An' when we got the wheel put on, away we went again.\\*
Them that had their noses broke had to go back ower hyem.\\
Some went to the dispensary, an' some to Dr Gibbs,\\*
An' some went to the infirmary to mend their broken ribs.\\!

An' when we got to {\sc Paradise}, there was bonnie gam's begun.\\*
There was four \& 20 on the bus, man. How they danced \& sung!\\
They called on me to sing a song, and I sang them ``Paddy Fagan''.\\*
An' I danced a jig an' I swung me twig on the day we went to {\sc Blaydon}.\\!

We flew across the chain bridge right into {\sc Blaydon} toon.\\*
The bellman he was callin' there. They call him \emph{Jacky Broon}.\\
I saw 'im talkin' to some chaps, an' he was them persuadin'\\*
To gan an' see \emph{Geordie Ridley}'s show at the {\sc Mechanics' Hall} in {\sc Blaydon}.\\!

The rain it poured al' the day an' made the ground quite muddy.\\*
\emph{Coffee John} 'ad a white hat on, an' he yelled, `We stole a cuddy!'\\
There was spice stalls \& monkey stalls an' old wives sellin' ciders.\\*
An' a chap wi' a ha'penny round about shoutin, `Now, me lads, for riders!'\\!

{\it O lads! You should've seen us gannin!\\*
Passin' the folks along the road, just as they were stan'in':\\
Al' the lads \& lasses there, an' al' wi' smilin' faces,\\*
Gannin along the {\sc Scotswood Road} to see the {\sc Blaydon} races.}
\end{verse}
\end{document}
