\documentclass[MAIN]{subfiles}

\begin{document}

\begin{center}
\emph{Tune: Smith of Bristol}
\end{center}

\bigskip

\settowidth{\versewidth}{With a noble crew of cutthroats, he used to scour the sea,}
\begin{verse}[\versewidth]
\emph{Smith} was a {\sc Bristol} man and a rare old sort was he.\footnotetext{`Smith of Bristol', Anon. \cite{dublinersa}. The origins of this song are obscure, but it may have been written in Ireland, and perhaps was originally intended as a parody, similar to ``The Auld Orange Flute''.}\\*
\vin {\it With his cutlass \& his pistol, heave your ho!}
With a noble crew of cutthroats, he used to scour the sea,\\
A-plundering and a-robbing high \& low.\\
He swore 'twas no concernin',\\
He didn't give a herrin',\\
About right or wrong or any holy show.\\
He swore that grabbin' booty\\
Was Britain's foremost duty,\\*
Wherever she could get it! Heave your ho!\\!

{\it Heave your ho! Heave your ho! Heave your ho!\\*
He swore that grabbin' booty\\
Was Britain's foremost duty,\\*
Wherever she could get it! Heave your ho!\\!}

\emph{Smith} had a noble soul and lofty was his pride.\\*
He'd watch his beaten foemen jump out into the tide,\\
All ye beggers who had nowhere else to go.\\
And hanging from his lanyards\\
Was portuguese and spaniards;\\
And beatmen frenchmen jumping to \& fro,\\
Right along the blazin' story,\\
Shall illumine England's glory:\\*
Pirate \emph{Smith} of {\sc Bristol}, heave your ho!\\!

But accidents can happen even to heroes such as he.\\*
\vin {\it With his cutlass \& his pistol, heave your ho!}\\
He was standing on his capstan as happy as could be,\\
Hoping soon to have another prize in tow,\\
When a whistling spanish bullet\\
Came and caught him in his gullet,\\
And very sad to say it laid him low.\\
He was only 97\\
When his soul had gone to heaven\\*
To rest on \emph{Nelson}'s bosom. Heave your ho!\\!

{\it Heave your ho! Heave your ho! Heave your ho!\\*
He swore that grabbin' booty\\
Was Britain's foremost duty,\\*
Wherever she could get it! Heave your ho!}

\end{verse}
\end{document}
